Modal Verification Patterns for Systems

### 2025-06-02 paper.tex
arxiv

\subsubsection{Virtualization}
One of the most common forms of abstraction provided by systems software is \emph{virtualization},
which abstracts the relationship between conceptual and physical computing resources. 
Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{Denning1970VM}).
Distributed language runtimes may virtualize addresses, even when processes may migrate across machines~\cite{jul1988fine}.
Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,Rosenblum1992LFS,bonwick2003zettabyte}.
Programs built on top of the corresponding systems software layer work logically at the level of these virtualized resources,
and it makes sense to specify the systems software directly in terms of those abstractions.
\paragraph{Access by Translation}
Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,Rosenblum1992LFS,rodeh2013btrfs}).
%Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer. 
B-trees~\cite{}, page tables~\cite{}, and related structures both behave like maps, when the corresponding physical resources exist as such
just in a different location.
Control over the lookup process (e.g, handling the case of a missing translation entry) allows for additional flexibility,
such as filling holes in sparse files, or demand paging (both from disk or lazily populating anonymous initially-zeroed mappings).
Although the realization of these maps may differ from a system to system based on the context (and sometimes hardware details),
they are semantically --- logically --- partial maps, worth treating as such in verification.

>>>>

\subsubsection{Virtualization}
One of the most common forms of abstraction provided by systems software is \emph{virtualization},
which abstracts the relationship between conceptual and physical computing resources. 
Operating System (OS) kernels virtualize memory locations and quantity (via virtual memory and paging~\cite{Denning1970VM}).
Distributed language runtimes may virtualize addresses, even when processes may migrate across machines~\cite{jul1988fine}.
Filesystems virtualize locations on disk~\cite{rodeh2013btrfs,hitz1994file,Rosenblum1992LFS,bonwick2003zettabyte}.
Programs built on top of the corresponding systems software layer work logically at the level of these virtualized resources,
and it makes sense to specify the systems software directly in terms of those abstractions.
\paragraph{Access by Translation}
Accessing virtualized resources via translation is a common way to virtualize notions of location (e.g.,
virtual memory addresses, inodes or object IDs instead of disk addresses~\cite{bonwick2003zettabyte,hitz1994file,Rosenblum1992LFS,rodeh2013btrfs}).
%Accessing the hardware resources in the presence of sharing means having more resource references (references to the virtualized resources) to access the \emph{mapped} resources that are physical and potentially fewer. 
B-trees, page tables, and related structures both behave like maps, when the corresponding physical resources exist as such
just in a different location.
Control over the lookup process (e.g, handling the case of a missing translation entry) allows for additional flexibility,
such as filling holes in sparse files, or demand paging (both from disk or lazily populating anonymous initially-zeroed mappings).
Although the realization of these maps may differ from a system to system based on the context (and sometimes hardware details),
they are semantically --- logically --- partial maps, worth treating as such in verification.
